---
title: "Homework 1"
author: Lunbei Hu
output: github_document
---

This is my solution to HW1.

```{r}
library(tidyverse)
```


## Problem 1

Created a data frame with specified elements.

```{r df}
prob1_df =
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
  )
```

Take the mean of each variable in my data frame.

```{r mean}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

I can take the mean of numbers and logical but not character or factor.

```{r as.numeric, results = "hide", warning=FALSE}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

The character vector cannot be returned as number, with the warning "NAs introduced by coercion." The factor vector can be returned as number 2 3 1 with our coercion. Since character and factor cannot be converted (or can be converted but with coercion) into number, we cannot get their means.

Here is the conversion for the logical variable to 1) numeric, 2) factor and 3) first factor then numeric, and the results are multiplied by the random sample.

```{r conversion}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```


## Problem 2

```{r import data}
data("penguins", package = "palmerpenguins")
```

**Here is the short description of the penguins dataset:**

* The dataset has `r nrow(penguins)` rows/observations and `r ncol(penguins)` columns/variables. 

* The eight variables are `r names(penguins)`.

* The three penguin species are `r levels(penguins$species)`, with frequencies `r filter(penguins, species == "Adelie") %>% nrow()`, `r filter(penguins, species == "Chinstrap") %>% nrow()`, and `r filter(penguins, species == "Gentoo") %>% nrow()` respectively.

* The three islands are `r levels(penguins$island)`, with frequencies `r filter(penguins, island == "Biscoe") %>% nrow()`,`r filter(penguins, island == "Dream") %>% nrow()`, and `r filter(penguins, island == "Torgersen") %>% nrow()` respectively.

* Four important descriptive variables are: bill length (mean `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`mm), bill depth (mean `r round(mean(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`mm), flipper length (mean `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)`mm), and body mass (mean `r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), digits = 2)`g).


Next I show a scatterplot of flipper length vs bill length - different color for different penguin species.

```{r yx_scatter}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

Export the scatterplot to my project directory.

```{r plot_export}
ggsave("bill_flipper.pdf", height = 4, width = 4)
```

